A search for high-mass black holes in data from the second LIGO observing run using a Bayesian ranking-statistic

### 2021-05-26 main.tex

\newcommand{\gstlal}{{\sc GstLAL}\xspace}
\newcommand{\cwb}{{\sc cWB}\xspace}
\newcommand{\pycbc}{{\sc \href{https://pycbc.org/}{{PyCBC}}}\xspace}

@@

% math keywords 
\newcommand{\fancytext}[1]{{\relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\mathcmd}[1]{{\sc \relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\bcr}{\mathcmd{\rho_\text{BCR}}}
\newcommand{\pycbcstat}{\mathcmd{\rho_\text{PC}}}
\newcommand{\snr}{\mathcmd{\rho}}
\newcommand{\psd}{\mathcmd{P(f)}}
\newcommand{\msun}{\mathcmd{\text{M}_\odot}}
\newcommand{\parameters}{\mathcmd{\vec{\theta}}}
\newcommand{\prior}{\mathcmd{\pi(\parameters)}}
\newcommand{\template}{\mathcmd{\mu(\parameters)}}
\newcommand{\fap}{\mathcmd{\text{FAP}}}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Theoretically, ground-based gravitational-wave detectors are sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detectors are plagued with short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect high-mass binary black hole mergers (with a total mass $>55$~\msun). We apply this technique to the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some stellar mass binary black holes unreported in GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($\text{M}_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($\text{M}_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations (M$_{BH}$-$\sigma$ and M$_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes, due to their small sphere of influence. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties due to the observational techniques utilized~\cite{Greene:2020:ARA&A}. Compact binaries coalesces (CBCs) can provide unambiguous gravitational wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). Ground-based gravitational wave detectors can probe the lower-end $10^{2} - 10^{3}\ \msun$ of the intermediate mass range.

Mergers involving either a binary system with at least one intermediate-mass component, or a system whose merger results in an intermediate-mass remnant will have large total masses $\text{M}_{T}$. As a binary's $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$, systems with large $\text{M}_{T}$ have very low merger frequencies $f < 100\ \text{Hz}$. Hence, ground based gravitational wave detectors ($\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($\sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}.

Because of high-total mass systems' short-durations in ground-based gravitational wave data steams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories are plagued with incoherent non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Some glitches, similar to signals from high total mass mergers, last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from high total mass mergers.

Although a significant fraction of the glitches can be removed by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to remove all the glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to account for more glitches while searching for high total mass CBC gravitational-waves is by utilizing a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. In this paper, we utilize a Bayesian method, called the Bayesian Coherence Odds-Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational-wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. The \bcr used in this study is a bootstrap-Bayesian odds computed using Bayesian evidences that describe the explicit probability of data under the hypothesis that the data contains coherent signals vs incoherent glitches, and empirically calculated prior-odds that describe how likely each hypotheses is. The high-mass candidate \bcr values are then used to calculate the probability that the candidate is of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.

@@

The standard framework to identify CBC gravitational-wave signals hidden in data is by quantifying the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent signal versus just noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

In a similar vein, this work utilizes Bayesian inference to calculate the Bayesian Coherence Odds-Ratio~\cite{BCR1}, \bcr, of high-mass candidates in LIGO's second observing run. We use the \bcr not as an odds-ratio but instead as a ranking statistic.

@@

Bayes theorem states that the posterior probability distribution $p(\parameters|d,\mathcal{H})$ for data $d$ and a vector of parameters \parameters that describe a model which quantifies a hypothesis $\mathcal{H}$, is given by
\begin{equation}
p(\parameters|d,\mathcal{H}) = \frac{\mathcal{L}(d|\parameters, \mathcal{H}) \ \pi(\parameters | \mathcal{H})}{\mathcal{Z}(d|\mathcal{H})}\ , 
\end{equation}
where $\mathcal{L}(d|\parameters, \mathcal{H})$ is the likelihood of the data given the parameters \parameters and the hypothesis, $\pi(\parameters | \mathcal{H})$ is the prior probability of the parameters, and finally,
\begin{equation}
    \mathcal{Z}(d|\mathcal{H}) = \int\limits_{\parameters} \ \mathcal{L}(d|\parameters,\mathcal{H}) \ \pi(\parameters | \mathcal{H}) \ d\parameters\ ,
\end{equation} is the likelihood after marginalizing over the parameters \parameters.  To compare two hypotheses $\mathcal{H}_A$ and $\mathcal{H}_B$ with the Bayes theorem one can calculate an odds-ratio
\begin{equation}
    \mathcal{O}^A_B = \frac{\mathcal{Z}^A\ \pi(\parameters | \mathcal{H_A})}{\mathcal{Z}^B\ \pi(\parameters | \mathcal{H_B})}\ ,
\end{equation}
where  $\{\pi(\parameters | \mathcal{H_A}), \pi(\parameters | \mathcal{H_B})\}$ are the prior-odds for each hypothesis and $\{\mathcal{Z}^A, \mathcal{Z}^B\}$ are shorthand for the evidences $\{\mathcal{Z}(d|\mathcal{H}_A), \mathcal{Z}(d|\mathcal{H}_B)\}$. The odds-ratio can quantify which of the two hypotheses is more likely. For example, if $\mathcal{O}^A_B >> 1$, then the odds are in favor of the $\mathcal{H}_A$ hypotheses.

The \bcr is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{P^S Z^S}{\prod\limits^D_{i=1} \ [P^G Z^G_i + (1-P^G)Z^N_i]}\ ,
\end{equation}
where $P^S$ and $P^G$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $P^S=P(\mathcal{H}_S)/P(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $P^G=P(\mathcal{H}_G| \mathcal{H}_I)$, the probability of obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

@@

\avi{
The pastro calculation may need some more discussion. Taking pastro = 1-FAP as identifying a real signal can be hugely problematic: \url{https://en.wikipedia.org/wiki/Misuse_of_p-values}, even though its ok in this case. Maybe the following papers have something that can help motivate this ~\cite{Farr:2015:PhRvD, Gaebel:2019:MNRAS,Galaudage:2020:PhRvD}
}

\subsection{Data for Analysis}
The LIGO Scientific collaboration operates several search pipelines that scan for gravitational-waves from compact binary mergers such as \gstlal~\cite{sachdev2019gstlal}, \texttt{MBTAOnline}~\cite{MBTA}, \texttt{SPIIR}~\cite{spiir} and \pycbc~\cite{pycbc_og4}. The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. The \pycbcstat ranking-statistic is akin to the matched-filter signal-to-noise ratio \snr. However, unlike \snr, \pycbcstat includes candidate signal's intrinsic and extrinsic properties and other information that feeds into determining if the signal can have astrophysical origins~\cite{pycbc_og6}. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}.

When \pycbc observes a trigger between detectors with coincident parameters and a time of arrival difference less than the gravitational-wave travel time between detectors, the trigger is labeled a \textit{candidate event trigger}, a trigger that may be from astrophysical origins~\cite{pycbc_og1}. To test the pipeline's sensitivity \pycbc also conducts searches for \textit{simulated triggers}, artificial triggers manufactured by injecting signals into the detector data. Finally, to quantify the statistical significance of candidate triggers, \pycbc artificially constructs \textit{background triggers} to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. The background trigger's \pycbcstat distribution is used to calculate the candidate trigger's significance,  using null-hypothesis significance testing, under the assumption that all candidate event triggers are due to noise.

Our work demonstrates that the \bcr can be used in the same way as \pycbcstat to measure candidate triggers' statistical significance.  The \bcr can be a powerful ranking statistic as it incorporates information of not only all possible binary black hole systems that might have merged to produce the trigger but also the various incoherent glitches that might cause a false-detection.

@@

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalizes over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}). While this marginalized likelihood reduces the run time without introducing errors to our evidence evaluation, it does not generates samples for the marginalized parameters. However, these parameter samples can be calculated as a post-processing step~\cite{intro_to_gw_bayes}.

>>>>

\newcommand{\gstlal}{{\sc GstLAL}\xspace}
\newcommand{\cwb}{{\sc cWB}\xspace}
\newcommand{\spiir}{{\sc SPIIR}\xspace}
\newcommand{\mbta}{{\sc MBTAOnline}\xspace}
\newcommand{\pycbc}{{\sc \href{https://pycbc.org/}{{PyCBC}}}\xspace}

@@

% math keywords 
\newcommand{\fancytext}[1]{{\relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\mathcmd}[1]{{\sc \relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\bcr}{\mathcmd{\rho_\text{BCR}}}
\newcommand{\pycbcstat}{\mathcmd{\rho_\text{PC}}}
\newcommand{\snr}{\mathcmd{\rho}}
\newcommand{\psd}{\mathcmd{P(f)}}
\newcommand{\msun}{\mathcmd{\text{M}_\odot}}
\newcommand{\mtot}{\mathcmd{M}}
\newcommand{\parameters}{\mathcmd{\vec{\theta}}}
\newcommand{\prior}{\mathcmd{\pi(\parameters)}}
\newcommand{\template}{\mathcmd{\mu(\parameters)}}
\newcommand{\fap}{\mathcmd{\text{FAP}}}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detectors are plagued with short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes, making their detection difficult. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect high-mass binary black hole mergers (with a total mass $>55$~\msun). We apply this technique to the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some stellar mass binary black holes unreported in GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the `intermediate-mass' range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties due to the observational techniques utilized~\cite{Greene:2020:ARA&A}.

Compact binaries coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). Ground-based gravitational wave detectors can probe the lower-end $10^{2} - 10^{3}\ \msun$ of the intermediate mass range. 
 As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$, systems with large $\text{M}_{T}$ have very low merger frequencies $f < 100\ \text{Hz}$. Hence, ground based gravitational wave detectors ($\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($\sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}.

Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detector data steams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories are plagued with non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Some glitches, similar to signals from high total mass mergers, last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from high total mass mergers~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be removed by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to remove all the glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to account for more glitches while searching for high total mass CBC gravitational-waves is by utilizing a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. In this paper, we utilize a Bayesian method, called the Bayesian Coherence Odds-Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational-wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. The \bcr used in this study is a bootstrap-Bayesian odds computed using Bayesian evidences that describe the explicit probability of data under the hypothesis that the data contains coherent signals vs incoherent glitches, and empirically calculated prior-odds that describe how likely each hypotheses is. \rs{Need to add another couple of sentences describing what is unique about this work. For instance, we study both foreground \textit{and} background events (which is the first time anyone's done a large-scale Bayesian analysis on background data). Suggest using this format: ``We highlight a number of innovations unique to this work. First...''} The high-mass candidate \bcr values are then used to calculate the probability that the candidate is of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.

@@

The standard framework to identify CBC gravitational-wave signals in data is by quantifying the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent signal versus just noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal vs. instrumental glitches or noise~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies and populations~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian ``evidence'': a marginalized likelihood. The evidence is the correct likelihood of having obtained the data given a hypothesis. However, due to the high computational cost to calculate the evidence, it is not explicitly used in current matched filter searches. Here, we explore a middle-ground hybrid between frequentist matched filtered searches and Bayesian odds. We compute Bayesian evidences for foreground and background data under the assumption that they either contain coherent gravitational-wave signals, noise, or glitches. However, instead of computing true Bayesian odds, we use the evidences to form a bootstrapped distribution for the odds of foreground and background events to form a frequentist ranking statistic.

@@

Bayes theorem states that the posterior probability distribution $p(\parameters|d,\mathcal{H})$ for data $d$ and a vector of parameters \parameters that describe a model which quantifies a hypothesis $\mathcal{H}$, is given by
\begin{equation}
p(\parameters|d,\mathcal{H}) = \frac{\mathcal{L}(d|\parameters, \mathcal{H}) \ \pi(\parameters | \mathcal{H})}{\mathcal{Z}(d|\mathcal{H})}\ , 
\end{equation}
where $\mathcal{L}(d|\parameters, \mathcal{H})$ is the likelihood of the data given the parameters \parameters and the hypothesis, $\pi(\parameters | \mathcal{H})$ is the prior probability of the parameters, and finally,
\begin{equation}
    \mathcal{Z}(d|\mathcal{H}) = \int\limits_{\parameters} \ \mathcal{L}(d|\parameters,\mathcal{H}) \ \pi(\parameters | \mathcal{H}) \ d\parameters\ ,
\end{equation} is the likelihood after marginalizing over the parameters \parameters.  To compare two hypotheses $\mathcal{H}_A$ and $\mathcal{H}_B$ with the Bayes theorem one can calculate an odds-ratio
\begin{equation}
    \mathcal{O}^A_B = \frac{\mathcal{Z}^A\ \pi^A}{\mathcal{Z}^B\ \pi^B}\ ,
\end{equation}
where  $\{\pi^A, \pi^B\}$ are the prior-odds for each hypothesis and $\{\mathcal{Z}^A, \mathcal{Z}^B\}$ are shorthand for the evidences $\{\mathcal{Z}(d|\mathcal{H}_A), \mathcal{Z}(d|\mathcal{H}_B)\}$. The odds-ratio can quantify which of the two hypotheses is more likely. For example, if $\mathcal{O}^A_B >> 1$, then the odds are in favor of the $\mathcal{H}_A$ hypotheses.

The \bcr is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\mathcal{O}_{\mathrm{BCR}} = \frac{P^S Z^S}{\prod\limits^D_{i=1} \ [P^G Z^G_i + (1-P^G)Z^N_i]}\ ,
\end{equation}

where $P^S$ and $P^G$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $P^S=P(\mathcal{H}_S)/P(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $P^G=P(\mathcal{H}_G| \mathcal{H}_I)$, the probability of obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

\begin{equation}
\label{eq:bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ ,
\end{equation}

\avi{Ranking statistic \bcr is like an estimator of what $\mathcal{O}_{\mathrm{BCR}}$ is}

@@

\avi{Eric, can you take a look at this section about the FAP p-astro?}
\avi{
The pastro calculation may need some more discussion. Taking pastro = 1-FAP as identifying a real signal can be hugely problematic: \url{https://en.wikipedia.org/wiki/Misuse_of_p-values}, even though its ok in this case. Maybe the following papers have something that can help motivate this ~\cite{Farr:2015:PhRvD, Gaebel:2019:MNRAS,Galaudage:2020:PhRvD}
}

\subsection{Data for Analysis}

\rs{Rory: Suggest a bit more preable here, something like ``To build a ranking statistic from $\rho$ we need to compute its distribution for foreground and background events. The number of background events in a typical search is very large, however we can manage the number of background triggers by focusing on a specific mass range. Background triggers are generated by search pipelines...}

The LIGO Scientific collaboration operates several search pipelines that scan for gravitational-waves from compact binary mergers such as \gstlal~\cite{sachdev2019gstlal}, \mbta~\cite{MBTA}, \spiir~\cite{spiir} and \pycbc~\cite{pycbc_og4}. The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. The \pycbcstat ranking-statistic is akin to the matched-filter signal-to-noise ratio \snr. However, unlike \snr, \pycbcstat includes candidate signal's intrinsic and extrinsic properties and other information that feeds into determining if the signal can have astrophysical origins~\cite{pycbc_og6}. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}.

\pycbc produces three categories of triggers: 
\begin{itemize}
    \item \textit{Candidate event trigger}: a trigger observed with coincident parameters amongst a network of detectors. 
    \item \textit{Simulated trigger}: a trigger detected from an artificially manufactured signal injected into detector data.  
    \item \textit{Background trigger}: a trigger obtained from incoherent signal-free background data (data manufactured by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}). 
\end{itemize}

% \rs{Rory: suggest cutting most or all of this paragraph. We don't need to explain how pycbc works in detail. A reference to the relevant paper is sufficient. The key point is that a pipeline produces them for us} When \pycbc observes a trigger between detectors with coincident parameters and a time of arrival difference less than the gravitational-wave travel time between detectors, the trigger is labeled a \textit{candidate event trigger}, a trigger that may be from astrophysical origins~\cite{pycbc_og1}. To test the pipeline's sensitivity \pycbc also conducts searches for \textit{simulated triggers}, artificial triggers manufactured by injecting signals into the detector data. Finally, to quantify the statistical significance of candidate triggers, \pycbc artificially constructs \textit{background triggers} to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. The background trigger's \pycbcstat distribution is used to calculate the candidate trigger's significance,  using null-hypothesis significance testing, under the assumption that all candidate event triggers are due to noise.

Our work demonstrates that the \bcr can be used measure candidate triggers' statistical significance.  The \bcr can be a powerful ranking statistic as it incorporates information of not only all possible binary black hole systems that might have merged to produce the trigger but also the various incoherent glitches that might cause a false-detection.

@@

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalizes over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}).

We set the priors $\pi(\parameters|\mathcal{H}_S)$ and $\pi(\parameters|\mathcal{H}_G)$ to be identical which reflects our ignorance of the distribution of the population properties of signals and signal-like glitches. These priors restrict signals with mass parameters in the ranges presented in Table~\ref{tab:parameters}. The spins are aligned over a uniform range for the dimensionless spin magnitude from $\left[0,1\right]$. The luminosity distance prior assigns probability uniformly in comoving volume, with an upper cutoff of $5\ \text{Gpc}$. The full list of the priors, along with their shapes, limits and boundary conditions are documented in Table~\ref{tab:priors}.
